---
title: "STAT 331 Portfolio"
author: "Leticia Mezzetti"
format: 
  html: 
    self-contained: true
    theme: 
      light: minty
      dark: minty
layout: margin-left
editor: visual
execute: 
  eval: false
  echo: true
---

[**My Grade:**]{.underline} I believe my grade equivalent to course work evidenced below to be an \_\_.

[**Learning Objective Evidence:**]{.underline} In the code chunks below, provide code from a Lab or Challenge assignment where you believe you have demonstrated proficiency with the specified learning target. Be sure to specify **where** the code came from (e.g., Lab 4 Question 2).

## Working with Data

**WD-1: I can import data from a *variety* of formats (e.g., csv, xlsx, txt, etc.).**

-   `csv`

```{r wd-1-csv}
# Taken from Lab 2, Question 1
surveys <- read_csv(here("Labs", "Lab 2", "surveys.csv"))
```

-   `xlsx`

```{r wd-1-xlsx}

```

-   `txt`

```{r wd-1-txt}

```

**WD-2: I can select necessary columns from a dataset.**

```{r wd-2}

```

**WD-3: I can filter rows from a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   numeric

```{r wd-3-numeric}

```

-   character -- specifically a string

```{r wd-3-string}

```

-   factor

```{r wd-3-factor}

```

-   date

```{r wd-3-date}

```

**WD-4: I can modify existing variables and create new variables in a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   numeric

```{r wd-4-numeric}

```

-   character -- specifically a string

```{r wd-4-string}

```

-   factor

```{r wd-4-factor}

```

-   date

```{r wd-4-date}

```

**WD-5: I can use mutating joins to combine multiple dataframes.**

-   `left_join()`

```{r wd-5-left}

```

-   `right_join()`

```{r wd-5-right}

```

-   `inner_join()`

```{r wd-5-inner}

```

-   `full_join()`

```{r wd-5-full}

```

**WD-6: I can use filtering joins to filter rows from a dataframe.**

-   `semi_join()`

```{r wd-6-semi}

```

-   `anti_join()`

```{r wd-6-anti}

```

**WD-7: I can pivot dataframes from long to wide and visa versa**

-   `pivot_longer()`

```{r wd-7-long}

```

-   `pivot_wider()`

```{r wd-7-wide}

```

## Reproducibility

**R-1: I can create professional looking, reproducible analyses using RStudio projects, Quarto documents, and the here package.**

I've done this in the following provided assignments:

**R-2: I can write well documented and tidy code.**

-   Example 1

```{r r-2-1}

```

-   Example 2

```{r r-2-2}

```

**R-3: I can write robust programs that are resistant to changes in inputs.**

-   Example 1

```{r r-3-1}

```

-   Example 2

```{r r-3-2}

```

## Data Visualization & Summarization

**DVS-1: I can create visualizations for a *variety* of variable types (e.g., numeric, character, factor, date)**

-   numeric variables

```{r dvs-1-num}

```

-   numeric variables and categorical variables

```{r dvs-2-num-cat}

```

-   categorical variables

```{r dvs-2-cat}

```

-   dates

```{r dvs-2-date}

```

**DVS-2: I use plot modifications to make my visualization clear to the reader.**

-   Example 1

```{r dvs-2-1}

```

-   Example 2

```{r dvs-2-2}

```

**DVS-3: I show creativity in my visualizations**

-   Example 1

```{r dvs-3-1}
# Taken from Lab 2, Question 14
ggplot(data = surveys,
        mapping = aes(x = species,
                      y = weight)) +
          geom_jitter(alpha = 0.5,
                      color = "cadetblue") +
            geom_boxplot(outlier.shape = NA) +
  labs(x = "Species of Rodents",
       y = "Weight (g)",
       title = "Relationship between Species of Rodents and Weight (g)") +
  guides(x = guide_axis(angle = 90))

# Axis Label Orientation: https://ggplot2.tidyverse.org/reference/guide_axis.html
```

-   Example 2

```{r dvs-3-2}
# Taken from Challenge 2, Question 2
cdPalette_grey <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
                    "#0072B2", "#D55E00", "#CC79A7")


ggplot(data = surveys,
        aes(x = weight, y = species, color = genus)) +
            geom_boxplot() +
  scale_color_manual(values = cdPalette_grey) +
  labs(x = "Weight (g)",
       y = "Species",
       color = "Genus of Organism") +
  annotate("text", y = 1, x = 400, label = "Neotoma") +
  annotate("text", y = 2, x = 200, label = "Chaetodipus") +
  annotate("text", y = 3, x = 300, label = "Peromyscus") +
  annotate("text", y = 4, x = 200, label = "Perognathus") +
  annotate("text", y = 5, x = 350, label = "Reithrodontomys") + 
  annotate("text", y = 6, x = 300, label = "Sigmodon") +
  annotate("text", y = 7, x = 300, label = "Onychomys") +
  annotate("text", y = 8, x = 300, label = "Peromyscus") +
  annotate("text", y = 9, x = 350, label = "Reithrodontomys") +
  annotate("text", y = 10, x = 350, label = "Dipodomys") +
  annotate("text", y = 11, x = 350, label = "Dipodomys") +
  annotate("text", y = 12, x = 350, label = "Chaetodipus") +
  annotate("text", y = 13, x = 350, label = "Dipodomys") +
  annotate("text", y = 14, x = 300, label = "Onychomys") +
  xlim(0, 600) + 
  theme_dark() +
  theme(legend.position = "none")
```

**DVS-4: I can calculate numerical summaries of variables.**

-   Example 1

```{r dvs-4-1}

```

-   Example 2

```{r dvs-4-2}

```

**DVS-5: I can find summaries of variables across multiple groups.**

-   Example 1

```{r dvs-5-1}

```

-   Example 2

```{r dvs-5-2}

```

**DVS-6: I can create tables which make my summaries clear to the reader.**

-   Example 1

```{r dvs-6-1}

```

-   Example 2

```{r dvs-6-2}

```

**DVS-7: I show creativity in my tables.**

-   Example 1

```{r dvs-7-1}

```

-   Example 2

```{r dvs-7-2}

```

## Program Efficiency

**PE-1: I can write concise code which does not repeat itself.**

-   using a single function call

```{r pe-1-one-call}

```

-   `across()`

```{r pe-1-across}

```

-   `map()` functions

```{r pe-1-map-1}

```

**PE-2: I can write functions to reduce repetition in my code.**

-   Example 1

```{r pe2-1}

```

-   Example 2

```{r pe2-2}

```

**PE-3:I can use iteration to reduce repetition in my code.**

-   `across()`

```{r pe-3-across}

```

-   `map()` functions (Provide 2 Examples)

```{r pe-3-map-1}

```

```{r pe-3-map-2}

```

**PE-4: I can use modern tools when carrying out my analysis.**

-   Example 1

```{r pe-4-1}

```

-   Example 2

```{r pe-4-2}

```

## Data Simulation & Modeling

**DSM-1: I can simulate data from a *variety* of probability models.**

-   Example 1

```{r dsm-1-1}

```

-   Example 2

```{r dsm-1-2}

```

**DSM-2: I can fit a linear regression and extract necessary summary measures.**

-   Example 1

```{r dsm-2-1}

```

-   Example 2

```{r dsm-2-2}

```

## Revising My Thinking

<!-- How did you revise your thinking throughout the course? How did you revise your thinking on the code examples you have provided in your portfolio? -->

## Extending My Thinking

<!-- How did you extended your thinking throughout the course? How did you extend your thinking on the code examples you have provided in your portfolio? -->

## Peer Support & Collaboration

<!-- Include an image of feedback you gave that you are proud of (either in a peer review or in Discord) -->
