---
title: "Challenge 8: Styling the 12 Days of Christmas"
author: "Leticia Mezzetti"
format: 
  html:
    self-contained: true
    code-tools: true
    code-fold: true
    theme: 
      light: minty
      dark: minty
execute: 
  echo: true
  messages: false
  warning: false
---

Options 1 & 2

```{r packages}
#| include: false
library(dplyr)
library(glue)
library(stringr)
library(purrr)
library(english)
```

### Test Data Set

```{r}
xmas <- read.csv("https://www.dropbox.com/s/e584pryn8evm1gz/xmas.csv?dl=1")

```

## Step One: Plurals

Note that the gifts are listed in singular: for example, on day five the narrator receives "five golden rings", but the entry in the dataset for the gift on day five simply says "ring".

Using the skeleton of the `pluralize_gift()` function, complete the code so that the function takes a gift and returns the appropriate plural.

```{r pluralize-function}
# Function that takes a noun and makes it plural
# Arguments -- gift -- A string or vector of strings
# Return -- A string or vector of strings with the pluralized words
# For str_squish: https://www.tidyverse.org/blog/2018/02/stringr-1-3-0/#:~:text=str_squish()%20%2C%20which%20removes%20whitespace,strings%20into%20a%20single%20space.

pluralize_gift <- function(gift){

gift <-
  if (str_detect(string = gift, pattern = "y$")){
    gift <- str_replace(string = gift, pattern = "y$","ies")
    } else if (str_detect(string = gift, pattern = "oose$")) {
      gift <- str_replace(string = gift, pattern = "oose$", "eese")
      } else {
        gift <- str_replace(string = gift, pattern = "$", "s")
      }

return(gift)

}

```

### Test Your Function

Try your function out on the smaller and then larger gift dataset. Consider: is your function *vectorized*? It does not have to be, but you can try it out if you want!

```{r pluralize-test}
## This will work if your function is vectorized!
#pluralize_gift(xmas$Gift.Item)

## This will work if your function is not vectorized!
purrr::map_chr(xmas$Gift.Item, pluralize_gift)

```

## Step Two: Creating sentences

Write a function called `make_phrase()` that takes as input the necessary information, and returns a phrase. For example,

    make_phrase(num_word = "ten", 
                item = "lords", 
                verb = "a-leaping", 
                adjective = "", 
                location = "")

should return

    "ten lords a-leaping"

```{r phrase-function}
make_phrase <- function(day, num_word, item, verb, adjective, location) {
  
  ## Step 1: Replace NAs with blank strings
  verb <- str_replace_na(verb, "")
  num_word <- str_replace_na(num_word, "")
  adjective <- str_replace_na(adjective, "")
  location <- str_replace_na(location, "")

  ## Step 2: If the day is larger than 1, the items need pluralized! 
    item <- case_when(num_word == 'one' ~ item,
                    TRUE ~ pluralize_gift(item))

  ## Step 3: If the day is 1, you need to add an "a" or "an" before the gift 
   num_word <- case_when(num_word == 'one' &
                           str_detect(string = item,
                                      pattern = "^[aeiou]") ~ 'an',
                         num_word == 'one' & 
                           !str_detect(string = item,
                                       pattern = "^[aeiou]") ~ 'a',
                         TRUE ~ num_word)

  ## Step 4: Glue all of the pieces together to make a phrase! 
  check <- glue::glue("{num_word} {adjective} {item} {verb} {location}")

return(check)
}

# Used this resource to format: https://evoldyn.gitlab.io/evomics-2018/ref-sheets/R_strings.pdf

```

### Test Your Function

Try your function out on the `xmas` data, by making a new variable containing the daily phrases. Notice I've provided you with the starter code to use `pmap_chr()` to iterate through each row of the dataset to create a phrase.

::: callout-tip
Look up the help file for `pmap_chr()` -- what is specified in the `list()`? How does that translate into what your `make_phrase()` function needs as inputs?
:::

```{r}
xmas2 <- xmas |>
  mutate(num = as.character(english::english(Day)), 
    Full.Phrase = pmap_chr(
      list(day = Day,
           num_word = num,
           item = Gift.Item,
           verb = Verb,
           adjective = Adjective,
           location = Location),
      make_phrase))

```

# Singing the Song

Write a function called `sing_day()` that takes as input:

-   A dataset (input as a dataframe)

-   A number indicating which day to sing about (input as an integer)

-   The name of a column in the dataset that contains the phrases for each day (input as an **tidy** name)

For example,

    sing_day(xmas2, 2, Full.Phrase)

should return

    On the second day of Christmas, my true love sent to me:
    two turtle doves and
    a partridge in a pear tree.

```{r}
# Produces the string for one day of the song.
# Arguments -- dataset: A data frame containing information about gifts
# Arguments -- line: The number of the line for the day you want to sing about
# Arguments -- phrase_col: The variable name for the column in the dataset that contains the song phrases
# Return -- A string singing the line of the song with all gifts for the given day.
# More info on the english package: https://cran.r-project.org/web/packages/english/english.pdf
# More info on glue_collapse: https://glue.tidyverse.org/reference/glue_collapse.html

sing_day <- function(dataset, num, phrase_col){
  
  # Step 1: Setup the intro line
  # Hint: You need to convert a number (e.g., 1) to a word (e.g., first)
  num_word <- english::ordinal(num)
  
  intro <- glue::glue("On the {num_word} day of Christmas, my true love gave to me")
  
  # Step 2: Sing the gift phrases

    phrases <- dataset |>
      pull( {{phrase_col}} )

    gifts <- phrases[num:1]
    combined <- glue::glue_collapse(gifts,
                                    sep = ",\n",
                                    last = ", and \n")
    
  ## put it together
    
   return (glue::glue("{intro} \n {combined} \n\n"))
}

```

### A Small Test

```{r}
sing_day(xmas2, num = 2, phrase_col = Full.Phrase)
```

## Using the Functions to Sing 12 Days of Christmas

```{r}
## Run previous to make it work
map_chr(1:12, ~sing_day(xmas2, .x, Full.Phrase)) |>
  str_c(collapse = "\n") |>
  cat()

# For examples on map_chr: https://rc2e.com/datatransformations, collapse: https://www.rdocumentation.org/packages/fstrings/versions/0.0.0.9000/topics/collapse, to concatenate/combine strings: https://www.statology.org/r-cat-function/

```

## 
