---
title: "Practice Activity 9"
author: "Leticia Mezzetti"
format: 
  html:
    self-contained: true
    code-tools: true
    code-fold: true
    theme: 
      light: minty
      dark: minty
execute: 
  echo: true
  messages: false
  warning: false
---

# **Practice Activity 9.1: Regression**

# **1 Data**

The dataset [`mystery_animal.csv`](https://atheobold.github.io/stat-331/9_simulation-regression/Practice%20Activity/mystery_animal.csv) contains the weights of a particular animal before and after a year of eating only roasted duck.

```{r}
library(tidyverse)
library(here)
library(dplyr)

mystery_animal <- read_csv(here("Week 9", "Practice Activities", "mystery_animal.csv"))
```

# **2 Linear Regression**

Fit a linear regression to determine if the duck diet is associated with the animal gaining weight, losing weight, or neither.

```{r}
mystery_lm <- mystery_animal |>
  lm(weight_after ~ weight_before, data = _)
```

# **3 Residuals**

Plot the residuals (observed value minus predicted value) of your linear model versus `weight_before`.

**Tip**

There are a few different ways to obtain your residuals. My favorite is the `augment()` function from the **broom** package. I like this option a lot because it gives you all of the information from your linear regression in a tidy [tibble]{.underline}! Each row has the explanatory and response values, the fitted / predicted value (`.fitted`), and the residual (`.resid`).

```{r}
broom::augment(mystery_lm)
```

# **4 Mystery Animal**

What animal do you see?

```{r}
mystery_lm |> 
  broom::augment() |> 
  ggplot(mapping = aes(y = .resid, x = .fitted)) +
  geom_point()
```
